---
title: "Intervals' method tutorial Part2"
author: "Thomas A. Püschel and Jordi Marcé-Nogué"
date: '2020-04-26T21:13:14-05:00'
output: rmarkdown::github_document
categories: R
tags:
- FEA
- Intervals
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse  = TRUE)
```

## Tutorial: How to use R to analyse data from FEA using the intervals’ method
# Part2: Convergence procedure to estimate an appropiate number of intervals

This script corresponds to the convergence procedure used to estimate an appropiate number of interval variables. We have provided the different files for the example using different numbers of intervals (i.e., N= 5, 10, 15, 25, 50, 75 and 100). When analysing your own data, you will need to create your .csv files with different number of intervals by changing the value of the ‘NIntervals’ object used in the previous tutorial (also available  as the “01-intervals-method-script” in  [figshare](https://figshare.com/articles/intervals-method-files_rar/12025866) and then run the script for each one of the cases.


# 1) Read all the files included in the convergence
 

```{r import data}
#SCALE = T to use the correlation matrix
#SCALE = F to use the variance-covariance matrix

intervals.data.5 <- read.csv("matrix-of-intervals-5.csv", row.names = 1, header = TRUE, sep = ",")
intervals.data.10 <- read.csv("matrix-of-intervals-10.csv", row.names = 1, header = TRUE, sep = ",")
intervals.data.15 <- read.csv("matrix-of-intervals-15.csv", row.names = 1, header = TRUE, sep = ",")
intervals.data.25 <- read.csv("matrix-of-intervals-25.csv", row.names = 1, header = TRUE, sep = ",")
intervals.data.50 <- read.csv("matrix-of-intervals-50.csv", row.names = 1, header = TRUE, sep = ",")
intervals.data.75 <- read.csv("matrix-of-intervals-75.csv", row.names = 1, header = TRUE, sep = ",")
intervals.data.100 <- read.csv("matrix-of-intervals-100.csv", row.names = 1, header = TRUE, sep = ",")
```

# 2) Compute PCAs for each one of the cases

```{r PCAs}
PCA.5 <- prcomp(intervals.data.5[, 1:5], scale = T)
PCA.10 <- prcomp(intervals.data.10[, 1:10], scale = T)
PCA.15 <- prcomp(intervals.data.15[, 1:15], scale = T)
PCA.25 <- prcomp(intervals.data.25[, 1:25], scale = T)
PCA.50 <- prcomp(intervals.data.50[, 1:50], scale = T)
PCA.75 <- prcomp(intervals.data.75[, 1:75], scale = T)
PCA.100 <- prcomp(intervals.data.100[, 1:100], scale = T)
```

# 3) Calculate te R-squared values for convergence procedure

PC1 convergence:
```{r}
Rvalues.pc1 <- c(
  summary(lm(PCA.10$x[, 1] ~ PCA.15$x[, 1]))$r.squared,
  summary(lm(PCA.15$x[, 1] ~ PCA.25$x[, 1]))$r.squared,
  summary(lm(PCA.25$x[, 1] ~ PCA.50$x[, 1]))$r.squared,
  summary(lm(PCA.50$x[, 1] ~ PCA.75$x[, 1]))$r.squared,
  summary(lm(PCA.75$x[, 1] ~ PCA.100$x[, 1]))$r.squared
)
```

PC2 convergence:
```{r}
Rvalues.pc2 <- c(
  summary(lm(PCA.10$x[, 2] ~ PCA.15$x[, 2]))$r.squared,
  summary(lm(PCA.15$x[, 2] ~ PCA.25$x[, 2]))$r.squared,
  summary(lm(PCA.25$x[, 2] ~ PCA.50$x[, 2]))$r.squared,
  summary(lm(PCA.50$x[, 2] ~ PCA.75$x[, 2]))$r.squared,
  summary(lm(PCA.75$x[, 2] ~ PCA.100$x[, 2]))$r.squared
)
```

Table with results: R-squared values

```{r}
data.pca <- data.frame(Rvalues.pc1, Rvalues.pc2)
names(data.pca) <- c("PC1", "PC2")
rownames(data.pca) <- c("PCA 10 vs. PCA 15", "PCA 15 vs. PCA 25", "PCA 25 vs. PCA 50", "PCA 50 vs. PCA 75", "PCA 75 vs. PCA 100")

rm(intervals.data.5, intervals.data.10, intervals.data.15, intervals.data.25, intervals.data.50, intervals.data.75, intervals.data.100)
```

# 4) Plot the PCAs and the R-squared values
```{r echo=FALSE, warning=FALSE}
plot(PCA.5$x[, 1], PCA.5$x[, 2], pch = 19, cex = 1.5, xlab = "", ylab = "", asp = T, main = "A) PCA 5 intervals", xlim = rev(range(PCA.5$x[, 1])), ylim = (range(PCA.5$x[, 2])))
plot(PCA.10$x[, 1], PCA.10$x[, 2], pch = 19, cex = 1.5, xlab = "", ylab = "", asp = T, main = "B) PCA 10 intervals", xlim = rev(range(PCA.10$x[, 1])), ylim = rev(range(PCA.10$x[, 2])))
plot(PCA.15$x[, 1], PCA.15$x[, 2], pch = 19, cex = 1.5, xlab = "", ylab = "", asp = T, main = "C) PCA 15 intervals", xlim = rev(range(PCA.15$x[, 1])), ylim = rev(range(PCA.15$x[, 2])))
plot(PCA.25$x[, 1], PCA.25$x[, 2], pch = 19, cex = 1.5, xlab = "", ylab = "", asp = T, main = "D) PCA 25 intervals", xlim = rev(range(PCA.25$x[, 1])), ylim = rev(range(PCA.25$x[, 2])))
plot(PCA.50$x[, 1], PCA.50$x[, 2], pch = 19, cex = 1.5, xlab = "", ylab = "", asp = T, main = "E) PCA 50 intervals", xlim = rev(range(PCA.50$x[, 1])), ylim = rev(range(PCA.50$x[, 2])))
plot(PCA.75$x[, 1], PCA.75$x[, 2], pch = 19, cex = 1.5, xlab = "", ylab = "", asp = T, main = "F) PCA 75 intervals", xlim = rev(range(PCA.75$x[, 1])), ylim = rev(range(PCA.75$x[, 2])))
plot(PCA.100$x[, 1], PCA.100$x[, 2], pch = 19, cex = 1.5, xlab = "", ylab = "", asp = T, main = "G) PCA 100 intervals", xlim = rev(range(PCA.100$x[, 1])), ylim = rev(range(PCA.100$x[, 2])))

plot('Pairwise PC correlations', 'R-squared values', main = "F) Convergence procedure", ylim = c(0.95, 1), xlim = c(1, 5))
points(Rvalues.pc1, col = c("tomato"), pch = 19, cex = 1.5)
points(Rvalues.pc2, col = c("salmon1"), pch = 19, cex = 1.5)
lines(Rvalues.pc1, lwd = 1, col = c("tomato"))
lines(Rvalues.pc2, lwd = 1, col = c("salmon1"))
abline(h = 1, lty = 3)
legend(0.97, legend = c('PC1','PC2'), pch = 19, col = c("salmon1", "tomato"), cex = 1)
```


